\documentclass[a4paper,11pt]{report}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{fullpage}
\usepackage{tikz}

\usetikzlibrary{graphs,graphs.standard}

\makeatletter
\pgfmathdeclarefunction{alpha}{1}{%
  \pgfmathint@{#1}%
  \edef\pgfmathresult{\pgffor@alpha{\pgfmathresult}}%
}

\usepackage{bussproofs}
\usepackage{mathpartir}
\usepackage{prooftrees}
\usepackage{color}

\newcommand*{\contract}[2]{contraction of $#1$ with $#2$}

\newcommand*{\MP}{Modus Ponens }
\newcommand*{\MPr}[2]{\MP of \ref{#1} and \ref{#2}}

\author{Sylvain Julmy}
\date{\today}

\setlength{\parindent}{0pt}

\begin{document}

\begin{center}
  \Large{
    Mathematical Methods for Computer Science 1\
    Spring 2017
  }
  \noindent\makebox[\linewidth]{\rule{\linewidth}{0.4pt}}

  Series 11
  \vspace*{1.4cm}

  Sylvain Julmy
  
  \noindent\makebox[\linewidth]{\rule{\linewidth}{0.4pt}}
\end{center}

\section*{\texttt{1}}

The grammar $G = (V,T,P,S)$ where

\begin{align*}
  V &= \{P\}\\
  T &= \{a,b\} \\
  P &= P \rightarrow a P a\ |\ b P b\ |\ \epsilon \\
  S &= P \\
\end{align*}

\section*{\texttt{2}}

\subsection*{\texttt{a)}}

This grammar generate the set of all possible balanced parenthesis string.

There are three possibilities :

\begin{enumerate}
\item $(S)$ : we pick an opened parenthesis, $S$ and a closed parenthesis, this
  is the only pick that consume the symbol of a string. So the opened
  parenthesis would always be closed when ``returning'' from $S$.
\item $SS$ : do not generate any symbols, but allow the grammar to accept
  sequence of balanced parenthesis terms.
\item $\epsilon$ : the empty string is balanced and allow to terminate the recursion.
\end{enumerate}

If we only pick $P = S \to (S)\ |\ \epsilon$, it is the set of any well balanced
term, by adding $SS$, we allow to reapeat this indefinitely.

\subsection*{\texttt{b)}}

This grammar generate the set of all possible operation with the operator $*$
and the symbol $a$. Terms of such an operator are always inside parenthesis.

The proof is by induction on the structure :

\paragraph{Base case : } in the base case, the accepted string is just the
single symbol $a$ or the product $(S * S)$ where $S$ is either a terminal or a
product $(S * S)$.

\paragraph{Induction : } at each step of the construction, we have to pick $S =
a$ or $S = (S * S)$. So the construction look likes an abstract syntax tree
(AST) for product expression.

A string accepted by this grammar is always well balanced, because only the
rules $S \to (S*S)$ generate parenthesis and an opened parenthesis is always closed.

\section*{\texttt{3}}

\begin{align*}
  V &= \{A,S,B\} \\
  T &= \{a,b\} \\
  P &= \{ \\
  S &\to ASB\ |\ \epsilon, \\
  A &\to aAS\ |\ a, \\
  B &\to SbS\ |\ A\ |\ bb \\
  \} \\
  S &= S
\end{align*}

\subsection*{\texttt{a)}}

$S$ is a nullable variable and $A$,$B$ are not nullable. Therefore, we obtain

\begin{align*}
  P &= \{\\
  S &\to ASB\ |\ AB ,\\
  A &\to aAS\ |\ aA\ |\ a ,\\
  B &\to SbS\ |\ bS\ |\ Sb\ |\ b\ |\ A\ |\ bb\\
  \}\\
\end{align*}

\subsection*{\texttt{b)}}

Unit pair : $(S,S)$, $(A,A)$ and $(B,B)$ are trivial unit pairs. Then $(B,A)$,
from $B \to A$, is a unit pair to. Therefore, we obtain

\begin{align*}
  P &= \{\\
  S &\to ASB\ |\ AB ,\\
  A &\to aAS\ |\ aA\ |\ a ,\\
  B &\to SbS\ |\ bS\ |\ Sb\ |\ b\ |\ aAS\ |\ aA\ |\ a |\ bb\\
  \}\\
\end{align*}

\subsection*{\texttt{c)}}

Each terminals is generating, therefore $a$ and $b$ are generating. There are no
production of $P$ that does not contains a non-generating symbol, therefore all
production are usefull.

\section*{\texttt{4}}

Assume $X \in V'' \cup T''$, we know that $X
\overset{*}{\underset{G}{\Rightarrow}} w$ for some $w \in T^*$ and every symbol
used in order to generate $w$ are generating, therefore $X
\overset{*}{\underset{G'}{\Rightarrow}}$.

Hence $X$ is not eliminated from $G''$, we know that $ \exists \alpha,\beta : S
\overset{*}{\underset{G'}{\Rightarrow}} \alpha X \beta$ and every symbol is
reachable, therefore $S \overset{*}{\underset{G''}{\Rightarrow}} \alpha X \beta$.

We know that every symbol in $\alpha X \beta$ is reachable and in $V' \cup
T'$, so they are generating in $G'$. The derivation of a string $xwy$ use only
symbols that are reachable from $S$ because they are reached by symbols in
$\alpha X \beta$.

Finally, $X$ can be any symbol in $G''$ and then $G''$ does not contains any
useless symbol.


\section*{\texttt{5}}

\subsection*{\texttt{a)}}

\begin{align*}
  G &= {V,T,P,S} \\
  V &= \{S\}\\
  T &= \{(,),\neg,\to,p,q\}\\
  P &= \{\\
  &S \to \neg S\ |\ (S \to S)\ |\ p\ |\ q \\
  \}\\
  S &= S\\
\end{align*}

$S$ occurs on the RHS of a production, therefore we create $S' \to S \in P$ :

\begin{align*}
  P &= \{\\
  &S' \to S, \\
  &S \to \neg S\ |\ (S \to S)\ |\ p\ |\ q \\
  \}\\
\end{align*}

We don't have to remove any $\epsilon$-production or unit productions, because
none of these are present.

Then, we transform RHS of productions which owns more than $2$ symbols :

\begin{align*}
  S &\to \neg S\ |\ (S \to S)\ |\ p\ |\ q \\
  \longrightarrow \\
  S &\to \neg S\ |\ (C_1     \ |\ p\ |\ q \\
  C_1 &\to S \to S ) \\
  \longrightarrow \\
  S &\to \neg S\ |\ (C_1     \ |\ p\ |\ q \\
  C_1 &\to S C_2 \\
  C_2 &\to \to S ) \\
  \longrightarrow \\
  S &\to \neg S\ |\ (C_1     \ |\ p\ |\ q \\
  C_1 &\to S C_2 \\
  C_2 &\to \to C_3 \\
  C_3 &\to S )
\end{align*}

Finally

\begin{align*}
  S' &\to S \\
  S &\to \neg S\ |\ (C_1     \ |\ p\ |\ q \\
  C_1 &\to S C_2 \\
  C_2 &\to \to C_3 \\
  C_3 &\to S ) \\
  \longrightarrow \\
  S' &\to S \\
  S &\to X S\ |\ Y C_1 \ |\ p\ |\ q \\
  C_1 &\to S C_2 \\
  C_2 &\to Z C_3 \\
  C_3 &\to S W \\
  W &\to ) \\
  X &\to \neg \\
  Y &\to ( \\
  Z &\to\ \to \\
\end{align*}

\subsection*{\texttt{b)}}

We have a string $w$ of length $n$, then we are going to take $n-1$ times a rule
of the form $A \to BC$ where $A$,$B$ and $C$ are non-terminal symbols, such that
we can construct a string with $n$ non-terminal symbols.

On each non-terminal symbol of said string of length $n$, we apply a rule of the
form $A \to a$ where $A$ is non-terminal and $a$ is terminal i.e. we apply $n$
rules. In total we have applied $n - 1 + = 2n-1$.

\end{document}
i